Comparing Apples and Oranges in Youth Depression Treatments?

### 2024-09-04 multilevel_metaregression/manuscript_with_multilevel_metan.qmd
added some text changes for revision

---
title: "Comparing Apples and Oranges in Youth Depression Treatments?"
subtitle: "A Quantitative Critique of the Evidence Base and Guidelines"
author:
  - name: Argyris Stringaris
    email: a.stringaris@ucl.ac.uk
    affiliations: 
        - id: ucl_1
          name: University College London
          department: Divisions of Psychiatry and Psychology and Language Science
          address: 1-19 Torrington Place
          city: London, UK
          state: 
          postal-code: WC1E 7HB
          country: United Kingdom
    attributes:
        corresponding: true
    note: 
  - name: Charlotte Burman
    email: charlotte.burman@ucl.ac.uk
    affiliations:
        - ref: ucl_1
    note:
  - name: Rudolf Uher
    email: uher@Dal.Ca
    affiliations:
        - id: dal
          name: Dalhousie University
          department: Department of Psychiatry
          address: 
          city: Nova Scotia
          state: 
          postal-code: 
          country: Canada
  - name: Dayna Bhudia
    email: dayna.bhudia@ucl.ac.uk
    affiliations:
        - ref: ucl_1
  - name: Despoina Miliou
    email: dgmiliou@outlook.com.gr
    affiliations:
        - ref: ucl_1
  - name: Ioannis-Marios Rokas
    email: <rokasgiannis93@gmail.com
    affiliations:
        - ref: ucl_1
  - name: Marinos Kyriakopoulos 
  - name: Lucy Foulkes
    email: lucy.foulkes@psych.ox.ac.uk
    affiliations:
        - id: OU
          name: University of Oxford
          department: Experimental Psychology
          address: Anna Watts Building, Radcliffe Observatory Quarter, Woodstock Road
          city: Oxford
          postal-code: OX2 6GG
          country: United Kingdom
    note:
  - name: Carmen Moreno
    email: cmoreno@hggm.es
    affiliations:
        - id: HGM
          name: Hospital Gregorio Marañón
          department: Department of Psychiatry
          address: 46 C. del Dr. Esquerdo
          city: Madrid
          state:
          postal-code: 28007
          country: Spain
  - name: Samuele Cortese
    email: samuele.cortese@soton.ac.uk
    affiliations:
        - id: soton
          name: University of Southampton
          department: Centre for Innovation in Mental Health (CIMH), School of Psychology
          address: Highfield Campus, Building 44
          city: Southampton
          state:
          postal-code: SO17 1BJ
          country: United Kingdom
  - name: Glyn Lewis
    email: glyn.lewis@ucl.ac.uk
    affiliations:
        - ref: ucl_1        
  - name: Georgina Krebs
    email: g.krebs@ucl.ac.uk
    affiliations:
        - ref: ucl_1
    note:
keywords: 
  - depression
  - adolescents
  - antidepressants
  - psychotherapy
date: last-modified
format:
  docx:
    reference-doc: custom-reference-doc.docx
    fig-width: 9
    fig-height: 7
    fig_caption: yes
    tbl_caption: yes
    numer-sections: true
editor: visual
bibliography: "emotion_concepts_paper.bib"
csl: "vancouver.csl"
font:
  family: "Calibri"
prefer-html: true
fig-format: png
fig-cap-location: top
---

@@

For adolescent depression, there are limited head-to-head trials of medication and psychotherapy, and hence recommendations are derived from indirect comparisons of treatment efficacy. The National Institute of Health and Care Excellence (NICE) guidelines for adolescent depression recommend psychotherapy (specifically cognitive behaviour therapy (CBT) or interpersonal therapy (IPT)) over medication in most cases [@niceDepressionChildrenYoung2019]. This is in keeping with two sources of evidence relating to child and adolescent depression: meta-analyses of medication randomised controlled trials (RCTs) that cast doubt on the efficacy of antidepressants, with the exception of fluoxetine [@ciprianiComparativeEfficacyTolerability2016]; and meta-analyses of psychotherapy RCTs that conclude psychotherapy to be efficacious [@cuijpersEffectsPsychologicalTreatments2021]. However, a recent network meta-analysis (NMA) [@zhouComparativeEfficacyAcceptability2020], an established method of comparing treatments using both direct and indirect (i.e. treatment A with treatment C, via studies that directly compare A with B and B with C) evidence, concluded that only fluoxetine alone and fluoxetine administered together with CBT were significantly more effective than medication control (i.e. pill placebo) or psychotherapy controls. The only head-to-head RCT comparing modalities found that fluoxetine, alone and in combination with CBT, was superior to pill placebo, though CBT alone was not [@marchFluoxetineCognitiveBehavioralTherapy2004]. Given this confusing evidence base, how should patients, carers, clinicians and policy makers make treatment decisions?

@@

For meta-analysis it is necessary to estimate a standard error of the SMD. This requires a correlation between the pre- and post-measures, a statistic typically not reported. To ensure that our results are not biased by misestimation, we simulated n=1000 datasets for different values (0.45 to 0.9) of this correlation and used these datasets in subsequent analyses (please refer to Supplement for full details).

@@

We conducted a series of sensitivity analyses where we excluded studies that 1) used waitlist as their control and 2) recruited participants with subclinical levels of depression. Next, we conducted two analyses where we included only trials that used the Children's Depression Rating Scale, Revised (CDRS-R) or the Hamilton Depression Rating Scale (HAM-D) as outcome instruments.

@@

@tbl-baseline_results summarises the results from each of the meta-analyses examining sample characteristics at baseline. The summary statistics are provided for each subgroup (i.e. medication and psychotherapy) and the p-value derives from the test for subgroup differences.

@@

```{r, warning=FALSE, message = FALSE, echo=FALSE, ft.align="left"}
#| label: tbl-baseline_results
#| tbl-cap: "Sample characteristics at baseline across medication and psychotherapy studies: Results for overall sample and sensitivity analyses"
#| tbl-cap-location: top

@@

On average, depression severity at baseline was significantly higher in medication trials compared to psychotherapy trials (see @tbl-baseline_results). When excluding RCTs that used waitlist as their control, baseline severity remained significantly higher in medication trials compared to psychotherapy trials. This difference did not reach statistical significance when excluding studies that recruited samples with sub-clinical depression.

```{r, warning=FALSE, message = FALSE, echo=FALSE}
# looking at CDRS studies only, this is with only 2 levels

@@

As can be seen in @tbl-baseline_results, mean age was `r round(results_met_age$TE_random[results_met_age$subgroup == 1][[2]],2)` (*SE* = `r round(results_met_age$seTE_random[results_met_age$subgroup == 1][[2]],2)`) across psychotherapy trials and `r round(results_met_age$TE_random[results_met_age$subgroup == 0][[2]],2)` (*SE* = `r round(results_met_age$seTE_random[results_met_age$subgroup == 0][[2]],2)`) across medication trials, with no significant between group differences. There were no significant differences in mean age between modalities on further sensitivity analyses.

@@

We applied metaregression to obtain the SMDs and confidence intervals of each of the four study arms. As seen in @fig-plot-means-all there were substantial differences between the four arms of the meta-analysis with striking differences between the medication and psychotherapy control arms. In particular, pill placebo had an SMD = `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "medication_control", ]$coef_means, 2)` (95% CI: `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "medication_control", ]$lower_ci, 2)` to `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "medication_control", ]$upper_ci, 2)`) whereas psychotherapy controls had an SMD = `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "psychotherapy_control", ]$coef_means, 2)` (95% CI: `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "psychotherapy_control", ]$lower_ci, 2)` to `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "psychotherapy_control", ]$upper_ci, 2)` ).

@@

df_no_sites <- readxl:: read_excel("for_despina_and_giannis_200224.xlsx")

@@

as_grouped_data(control_summary, groups = "Column") %>% 
  as_flextable(hide_grouplabel = TRUE ) %>% 
  bold(i = c(1, 4, 7, 10), bold = TRUE, part = "body") %>%  
  set_table_properties(width = 1) %>%  # Set table width 
  autofit() %>%  # Autofit column widths 
  align(j = 2:9, align = "center", part = "all") %>% 
  # add_header_row(values = "Comparing the intensity of the intervention between active and control arms of psychotherapy studies", colwidths = 4) %>% # This is a work around for the problems with rendering captions
  # add_header_row(values = "Table 5", colwidths = 4) %>% 
  bold(part = "header") %>%   # Bold the relevant header rows
  bg(bg = "#F2F2F2", part = "header") %>%   # Add light grey background to header row %>% 
  #italic(i = 2, italic = TRUE, part = "header") %>% 
  add_footer_row(values = as_paragraph("Abbreviations: K = number of studies, SD = standard deviation, CI = 95% confidence interval, df = degrees of freedom"), colwidths = 10) %>%
  font(fontname = "Calibri", part = "all") %>%   # Set font to Calibri
  border_remove() %>% # this is to remove borders around the caption that look strange
  hline(j = 1:4, part = "body", border = small_border) %>%  # Add light grey horizontal lines all  rows
  vline(j = c(1,4), part = "body", border = small_border) %>% 
  border_outer(border = small_border, part = "body" ) %>%  # Add outer borders
  hline(part = "header", border = big_border) %>% 
  #padding(i = 2, padding.bottom = 15, part = "header") %>% 
  paginate(init = TRUE) %>% 
  line_spacing(space = .7, part = "all") 
```

@@

Second, psychotherapy controls have moderate effect sizes (-0.5) whereas medication controls have very large effect sizes (-1.9). Our analysis could be critiqued as it compares within arm symptom change per trial, therefore effectively breaking randomisation. This criticism would apply if our aim were to draw inferences about the efficacy of each arm --- in which case preserving randomisation to balance confounders is critical. However, we note that our findings are largely in keeping with those of the NMA, which is designed to preserve the randomisation structure. Indeed, in Zhou et al. [@zhouComparativeEfficacyAcceptability2020] the estimates for psychotherapy controls, TAU and waitlist conditions favoured placebo (though CIs were broad because these were indirectly estimated), as did estimates for psychodynamic and behavioural therapy. CBT did not differentiate from placebo, a result that is likely heavily weighted by the results of their direct comparison in the TADS trial [@marchFluoxetineCognitiveBehavioralTherapy2004]. More importantly, we do not claim that these differences are genuinely due to efficacy differences; they may well be because people who attend psychotherapy and medication trials are different and therefore respond differently. In either case (difference in efficacy vs difference in trial participant profile), the substantial disparity in the response to control conditions is reason for concern about our ability to draw inferences from comparisons of modalities. This is particularly problematic as clinicians as well as policy makers often resort to between-group effect sizes to summarise findings, which our findings make obvious is misleading.

@@

We conducted a systematic search for medication studies published from 31 May 2015 up to 1 Jan 2021 (i.e. after the final search date of Cipriani et al.'s [@ciprianiComparativeEfficacyTolerability2016] review up to the final search date of Cuijpers et al's [@cuijpersEffectsPsychologicalTreatments2021] review). We searched PubMed, the Cochrane Central Register of Controlled Trials, Embase, Web of Science, CINAHL, PsycINFO and LiLACS for randomised controlled trials (RCTs) comparing any antidepressant with placebo in the treatment of children and adolescents with a primary diagnosis of major depressive disorder. We used the same search terms as Cipriani [@ciprianiComparativeEfficacyTolerability2016] with one additional search term to include only placebo-controlled trials (see below). We additionally applied filters to specify our date range, and to exclude reviews and non-human studies. We also searched clinical trial registers for published and unpublished studies however all RCTs meeting inclusion criteria had already been identified from the database search outlined above. Please see @fig-prisma for the PRISMA flow diagram.

@@

We drew upon RCTs included in two recent comprehensive meta-analyses with open data available for each medication and psychotherapy, and supplemented them with an updated systematic review. Please refer to these original meta-analyses for a detailed description of their search strategy and study selection criteria. Psychotherapy studies were drawn from a systematic review and meta-analysis of randomised trials comparing psychotherapy for youth depression against control conditions [@cuijpersEffectsPsychologicalTreatments2021] (dataset available [here](https://docs.metapsy.org/databases/depression-childadol-psyctr/)). Whilst Cuijpers et al. [@cuijpersEffectsPsychologicalTreatments2021] excluded studies for which the primary outcome variable could not be calculated due to missing data, we included these studies and performed the imputations outlined below. We also included studies which had data available for other variables in interest, including number of sites or baseline demographics; hence we have more psychotherapy studies included in this review compared to the original meta-analysis. Whilst the online database is regularly updated, we chose to exclude studies published after the final date of Cuijpers et al.'s [@cuijpersEffectsPsychologicalTreatments2021] literature search.

@@

where, $$\bar{X}$$ is the mean score for each study arm on the primary outcome questionnaire, and ${scale}_{min}$ and ${scale}_{max}$ are the minimum and maximum possible values of the scale in question, respectively. The standard deviation is calculated thus:

@@

$$θij​∼N(xij​β,τi2​)$$

{#eq-8}

@@

$$ Yij=xijβ+ui+vij+ϵij $$

{#eq-9}

@@

$$ ui​∼N(0,τ2) $$

{#eq-10} describes the deviation of each study from the overall mean effect size, and

$$vij∼N(0,τi2)$$ {#eq11}

@@

$$\epsilon_{ij} \sim \mathcal{N}(0, \sigma_{ij}^2)$$

{#eq-12}

@@

$$ \begin{aligned} Υ_{ij} &= \begin{cases} 0 & \text{MedControl:} \quad b_0 + u_i + v_{ij} + \epsilon_{ij}\\ 1 & \text{MedActive:} \quad b_0 + b_{1_{ij}} + u_i + v_{ij} + \epsilon_{ij} \\ 2 & \text{PsyActive:} \quad b_0 + b_{2_{ij}} + u_i + v_{ij} + \epsilon_{ij} \\ 3 & \text{PsyControl:} \quad b_0 + b_{3_{ij}} + u_i + v_{ij} + \epsilon_{ij} \\ \end{cases} \end{aligned} $$

{#eq-13} Here, the mean effect size for each level is the sum of $b_0$, the intercept for the reference category (medication control), with the coefficient for each level (e.g., $b_{3_{ij}}$ for psychotherapy controls). The variability between studies and arms within studies is captured by $u_i$ and $v_{ij}$, respectively. The confidence intervals for the means are constructed using the standard errors of the means, which account for the hierarchical structure of the data. Each coefficient represents the contrast between the reference category and each level. For example, $b_{3_{ij}}$ represents the contrast between psychotherapy and medication control arms. Inference on these contrasts is conducted using the following test statistic:

$$ z = \frac{\hat{\beta}}{\text{SE}(\hat{\beta})} $$

{#eq-14} This test allows us to assess the significance of the differences between treatment effects across study arms.

@@

We conducted a series of sensitivity analyses. For each of the meta-analyses we excluded studies that 1) used waitlist as their control and 2) recruited participants with subclinical levels of depression. Next, we conducted two analyses where we included only trials that used the Children's Depression Rating Scale, Revised (CDRS-R) or the Hamilton Depression Rating Scale (HAM-D) as outcome instruments.

@@

```{r, warning=FALSE, message = FALSE, echo=FALSE}
#| label: tbl-all-trials
#| tbl-cap: "Summary of included RCTs"
#| tbl-cap-location: top

@@

 #png("plot_means_multi_no_wl.png")
 #plot_means_no_wl
 #dev.off()

@@

# png("plot_means_multi_cdrs.png")
# plot_means_cdrs
# dev.off()

@@

# subtitle_text_hamd <- "Metanalytic estimates of within-group changes: HAM-D studies only"
plot_means_hamd <- plot_means_function(coef_and_se_means_hamd_study)
print(plot_means_hamd)

# png("plot_means_multi_cdrs.png")
# plot_means_cdrs
# dev.off()

# subtitle_text_hamd <- "Metanalytic estimates of within-group changes: HAM-D studies only"
plot_means_hamd <- plot_means_function(coef_and_se_means_hamd_study)
print(plot_means_hamd)

# png("plot_means_multi_cdrs.png")
# plot_means_cdrs
# dev.off()

@@

#| label: fig-plot-means-var2
#| fig-cap: "Meta-analytic estimates of within-group changes: studies with variance <0.02"
#| fig-cap-location: top

>>>>

---
title: "Comparing Apples and Oranges in Youth Depression Treatments?"
subtitle: "A Quantitative Critique of the Evidence Base and Guidelines"
author:
  - name: Argyris Stringaris
    email: a.stringaris@ucl.ac.uk
    affiliations: 
        - id: ucl_1
          name: University College London
          department: Divisions of Psychiatry and Psychology and Language Science
          address: 1-19 Torrington Place
          city: London, UK
          state: 
          postal-code: WC1E 7HB
          country: United Kingdom
    attributes:
        corresponding: true
    note: 
  - name: Charlotte Burman
    email: charlotte.burman@ucl.ac.uk
    affiliations:
        - ref: ucl_1
    note:
  - name: Raphaelle Delpech
    email: raphaelle.delpech.18@ucl.ac.uk
    affiliations:
        - ref: ucl_1
  - name: Rudolf Uher
    email: uher@Dal.Ca
    affiliations:
        - id: dal
          name: Dalhousie University
          department: Department of Psychiatry
          address: 
          city: Nova Scotia
          state: 
          postal-code: 
          country: Canada
  - name: Dayna Bhudia
    email: dayna.bhudia@ucl.ac.uk
    affiliations:
        - ref: ucl_1
  - name: Despoina Miliou
    email: dgmiliou@outlook.com.gr
    affiliations:
        - ref: ucl_1
  - name: Ioannis-Marios Rokas
    email: rokasgiannis93@gmail.com
    affiliations:
        - ref: ucl_1
  - name: Marinos Kyriakopoulos
  - name: Lucy Foulkes
    email: lucy.foulkes@psych.ox.ac.uk
    affiliations:
        - id: OU
          name: University of Oxford
          department: Experimental Psychology
          address: Anna Watts Building, Radcliffe Observatory Quarter, Woodstock Road
          city: Oxford
          postal-code: OX2 6GG
          country: United Kingdom
    note:
  - name: Carmen Moreno
    email: cmoreno@hggm.es
    affiliations:
        - id: HGM
          name: Hospital Gregorio Marañón
          department: Department of Psychiatry
          address: 46 C. del Dr. Esquerdo
          city: Madrid
          state:
          postal-code: 28007
          country: Spain
  - name: Samuele Cortese
    email: samuele.cortese@soton.ac.uk
    affiliations:
        - id: soton
          name: University of Southampton
          department: Centre for Innovation in Mental Health (CIMH), School of Psychology
          address: Highfield Campus, Building 44
          city: Southampton
          state:
          postal-code: SO17 1BJ
          country: United Kingdom
  - name: Glyn Lewis
    email: glyn.lewis@ucl.ac.uk
    affiliations:
        - ref: ucl_1        
  - name: Georgina Krebs
    email: g.krebs@ucl.ac.uk
    affiliations:
        - ref: ucl_1
    note:
keywords: 
  - depression
  - adolescents
  - antidepressants
  - psychotherapy
date: last-modified
format:
  docx:
    reference-doc: custom-reference-doc.docx
    fig-width: 9
    fig-height: 7
    fig_caption: yes
    tbl_caption: yes
    numer-sections: true
editor: visual
bibliography: "emotion_concepts_paper.bib"
csl: "vancouver.csl"
font:
  family: "Calibri"
prefer-html: true
fig-format: png
fig-cap-location: top
---

@@

For adolescent depression, there are limited head-to-head trials of medication and psychotherapy, and hence recommendations are derived from indirect comparisons of treatment efficacy. The National Institute of Health and Care Excellence (NICE) guidelines for adolescent depression recommend psychotherapy (specifically cognitive behaviour therapy (CBT) or interpersonal therapy (IPT)) over medication in most cases [@niceDepressionChildrenYoung2019]. This is in keeping with two sources of evidence relating to child and adolescent depression: meta-analyses of medication randomised controlled trials (RCTs) that cast doubt on the efficacy of antidepressants, with the exception of fluoxetine [@ciprianiComparativeEfficacyTolerability2016]; and meta-analyses of psychotherapy RCTs that conclude psychotherapy to be efficacious [@cuijpersEffectsPsychologicalTreatments2021]. However, a recent network meta-analysis (NMA) [@zhouComparativeEfficacyAcceptability2020], an established method of comparing treatments using both direct and indirect (i.e. treatment A with treatment C, via studies that directly compare A with B and B with C) evidence, concluded that only fluoxetine alone and fluoxetine administered together with CBT were significantly more effective than medication control (i.e. pill placebo) or psychotherapy controls. The only head-to-head RCT comparing modalities found that fluoxetine, alone and in combination with CBT, was superior to pill placebo, though CBT alone was not [@marchFluoxetineCognitiveBehavioralTherapy2004]. Also, the addition of psychotherapy to standard care did not improve outcomes (REF). Given this confusing evidence base, how should patients, carers, clinicians and policy makers make treatment decisions?

@@

For meta-analysis it is necessary to estimate a standard error of the SMD. This requires a correlation between the pre- and post-measures, a statistic typically not reported. To ensure that our results are not biased by misestimation (REF), we simulated n=1000 datasets for different values (0.45 to 0.9) of this correlation and used these datasets in subsequent analyses (please refer to Supplement for full details).

@@

We conducted a series of sensitivity analyses where we excluded studies that 1) used waitlist as their control and 2) recruited participants with subclinical levels of depression. Next, we conducted two analyses where we included only trials that used the Children's Depression Rating Scale, Revised (CDRS-R) or the Hamilton Depression Rating Scale (HAM-D) as outcome instruments. Additonally, we restricted two analyses to studies with variance below 0.02 and which reported SD at outcome respectively.

@@

tbl-baseline-results summarises the results from each of the meta-analyses examining sample characteristics at baseline. The summary statistics are provided for each subgroup (i.e. medication and psychotherapy) and the p-value derives from the test for subgroup differences.

@@

```{r, warning=FALSE, message = FALSE, echo=FALSE, ft.align="left"}
#| label: tbl-baseline-results
#| tbl-cap: "Sample characteristics at baseline across medication and psychotherapy studies: Results for overall sample and sensitivity analyses"
#| tbl-cap-location: top

@@

On average, depression severity at baseline was significantly higher in medication trials compared to psychotherapy trials (see @tbl-baseline-results). When excluding RCTs that used waitlist as their control, baseline severity remained significantly higher in medication trials compared to psychotherapy trials. This difference did not reach statistical significance when excluding studies that recruited samples with sub-clinical depression.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# looking at CDRS studies only, this is with only 2 levels

@@

As can be seen in tbl-baseline_results, mean age was `r round(results_met_age$TE_random[results_met_age$subgroup == 1][[2]],2)` (*SE* = `r round(results_met_age$seTE_random[results_met_age$subgroup == 1][[2]],2)`) across psychotherapy trials and `r round(results_met_age$TE_random[results_met_age$subgroup == 0][[2]],2)` (*SE* = `r round(results_met_age$seTE_random[results_met_age$subgroup == 0][[2]],2)`) across medication trials, with no significant between group differences. There were no significant differences in mean age between modalities on further sensitivity analyses.

@@

We applied metaregression to obtain the SMDs and confidence intervals of each of the four study arms. As seen in fig-plot-means-all there were substantial differences between the four arms of the meta-analysis with striking differences between the medication and psychotherapy control arms. In particular, pill placebo had an SMD = `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "medication_control", ]$coef_means, 2)` (95% CI: `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "medication_control", ]$lower_ci, 2)` to `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "medication_control", ]$upper_ci, 2)`) whereas psychotherapy controls had an SMD = `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "psychotherapy_control", ]$coef_means, 2)` (95% CI: `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "psychotherapy_control", ]$lower_ci, 2)` to `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "psychotherapy_control", ]$upper_ci, 2)` ).

@@

df_no_sites <- readxl:: read_excel(here("for_despina_and_giannis_200224.xlsx"))

@@

as_grouped_data(control_summary, groups = "Column") %>% 
  as_flextable(hide_grouplabel = TRUE ) %>% 
  bold(i = c(1, 4, 7, 10), bold = TRUE, part = "body") %>%  
  set_table_properties(width = 1) %>%  # Set table width 
  autofit() %>%  # Autofit column widths 
  align(j = 2:9, align = "center", part = "all") %>% 
  # add_header_row(values = "Comparing the intensity of the intervention between active and control arms of psychotherapy studies", colwidths = 4) %>% # This is a work around for the problems with rendering captions
  # add_header_row(values = "Table 5", colwidths = 4) %>% 
  bold(part = "header") %>%   # Bold the relevant header rows
  bg(bg = "#F2F2F2", part = "header") %>%   # Add light grey background to header row %>% 
  #italic(i = 2, italic = TRUE, part = "header") %>% 
  add_footer_row(values = as_paragraph("Abbreviations: K = number of studies, SD = standard deviation, CI = 95% confidence interval, df = degrees of freedom."), colwidths = 10) %>%
  font(fontname = "Calibri", part = "all") %>%   # Set font to Calibri
  border_remove() %>% # this is to remove borders around the caption that look strange
  hline(j = 1:4, part = "body", border = small_border) %>%  # Add light grey horizontal lines all  rows
  vline(j = c(1,4), part = "body", border = small_border) %>% 
  border_outer(border = small_border, part = "body" ) %>%  # Add outer borders
  hline(part = "header", border = big_border) %>% 
  #padding(i = 2, padding.bottom = 15, part = "header") %>% 
  paginate(init = TRUE) %>% 
  line_spacing(space = .7, part = "all") 
```

@@

Second, psychotherapy controls have moderate effect sizes (-0.5) whereas medication controls have very large effect sizes (-1.9). Our analysis could be critiqued as it compares within arm symptom change per trial, therefore effectively breaking randomisation. This criticism would apply if our aim were to draw inferences about the efficacy of each arm --- in which case preserving randomisation to balance confounders is critical. More importantly, we do not claim that these differences are genuinely due to efficacy differences; they may well be because people who attend psychotherapy and medication trials are different and therefore respond differently. In either case (difference in efficacy vs difference in trial participant profile), the substantial disparity in the response to control conditions is reason for concern about our ability to draw inferences from comparisons of modalities. This is particularly problematic as clinicians as well as policy makers often resort to between-group effect sizes to summarise findings, which our findings make obvious is misleading.

We note that our findings are largely in keeping with those of the NMA, which is designed to preserve the randomisation structure. Indeed, in Zhou et al. [@zhouComparativeEfficacyAcceptability2020] the estimates for psychotherapy controls, TAU and waitlist conditions favoured placebo (though CIs were broad because these were indirectly estimated), as did estimates for psychodynamic and behavioural therapy. CBT did not differentiate from placebo, a result that is likely heavily weighted by the results of their direct comparison in the TADS trial [@marchFluoxetineCognitiveBehavioralTherapy2004]. We note that the Zhou metanalysis, which represents an admirable effort to synthesise the literature, reports on issues that may affect transitivity with tests of incoherence showing significant differences.

@@

We conducted a systematic search for medication studies published from 31 May 2015 up to 1 Jan 2021 (i.e. after the final search date of Cipriani et al.'s [@ciprianiComparativeEfficacyTolerability2016] review up to the final search date of Cuijpers et al's [@cuijpersEffectsPsychologicalTreatments2021] review). We searched PubMed, the Cochrane Central Register of Controlled Trials, Embase, Web of Science, CINAHL, PsycINFO and LiLACS for randomised controlled trials (RCTs) comparing any antidepressant with placebo in the treatment of children and adolescents with a primary diagnosis of major depressive disorder. We used the same search terms as Cipriani [@ciprianiComparativeEfficacyTolerability2016] with one additional search term to include only placebo-controlled trials (see below). We additionally applied filters to specify our date range, and to exclude reviews and non-human studies. We also searched clinical trial registers for published and unpublished studies however all RCTs meeting inclusion criteria had already been identified from the database search outlined above. Please see fig-prisma for the PRISMA flow diagram.

@@

We drew upon RCTs included in two recent comprehensive meta-analyses with open data available for each medication and psychotherapy, and supplemented them with an updated systematic review. Please refer to these original meta-analyses for a detailed description of their search strategy and study selection criteria. Psychotherapy studies were drawn from a systematic review and meta-analysis of randomised trials comparing psychotherapy for youth depression against control conditions [@cuijpersEffectsPsychologicalTreatments2021] (dataset available at <https://docs.metapsy.org/databases/depression-childadol-psyctr/>). Whilst Cuijpers et al. [@cuijpersEffectsPsychologicalTreatments2021] excluded studies for which the primary outcome variable could not be calculated due to missing data, we included these studies and performed the imputations outlined below. We also included studies which had data available for other variables in interest, including number of sites or baseline demographics; hence we have more psychotherapy studies included in this review compared to the original meta-analysis. Whilst the online database is regularly updated, we chose to exclude studies published after the final date of Cuijpers et al.'s [@cuijpersEffectsPsychologicalTreatments2021] literature search. We also included three studies from Zhou et al. (REF) which were not covered in the previous meta-analysis whilst fitting the date range.

@@

where, $$\bar{X}$$is the mean score for each study arm on the primary outcome questionnaire, and ${scale}_{min}$ and ${scale}_{max}$ are the minimum and maximum possible values of the scale in question, respectively. The standard deviation is calculated thus:

@@

$$θij​∼N(xij​β,τi2​)$$ {#eq-8}

@@

$$ Yij=xijβ+ui+vij+ϵij $$ {#eq-9}

@@

$$ ui​∼N(0,τ2) $$ {#eq-10}

describes the deviation of each study from the overall mean effect size, and

$$vij∼N(0,τi2)$$ {#eq-11}

@@

$$\epsilon_{ij} \sim \mathcal{N}(0, \sigma_{ij}^2)$$ {#eq-12}

@@

$$ \begin{aligned} Υ_{ij} &= \begin{cases} 0 & \text{MedControl:} \quad b_0 + u_i + v_{ij} + \epsilon_{ij}\\ 1 & \text{MedActive:} \quad b_0 + b_{1_{ij}} + u_i + v_{ij} + \epsilon_{ij} \\ 2 & \text{PsyActive:} \quad b_0 + b_{2_{ij}} + u_i + v_{ij} + \epsilon_{ij} \\ 3 & \text{PsyControl:} \quad b_0 + b_{3_{ij}} + u_i + v_{ij} + \epsilon_{ij} \\ \end{cases} \end{aligned} $$ {#eq-13}

Here, the mean effect size for each level is the sum of $b_0$, the intercept for the reference category (medication control), with the coefficient for each level (e.g., $b_{3_{ij}}$ for psychotherapy controls). The variability between studies and arms within studies is captured by $u_i$ and $v_{ij}$, respectively. The confidence intervals for the means are constructed using the standard errors of the means, which account for the hierarchical structure of the data. Each coefficient represents the contrast between the reference category and each level. For example, $b_{3_{ij}}$ represents the contrast between psychotherapy and medication control arms. Inference on these contrasts is conducted using the following test statistic:

$$ z = \frac{\hat{\beta}}{\text{SE}(\hat{\beta})} $$ {#eq-14}

This test allows us to assess the significance of the differences between treatment effects across study arms.

@@

We conducted a series of sensitivity analyses. For each of the meta-analyses we excluded studies that 1) used waitlist as their control and 2) recruited participants with subclinical levels of depression. Next, we conducted two analyses where we included only trials that used the Children's Depression Rating Scale, Revised (CDRS-R) or the Hamilton Depression Rating Scale (HAM-D) as outcome instruments. Additonally, we restricted two analyses to studies with variance below 0.02 and which reported SD at outcome respectively.

@@

```{r echo=FALSE, message=FALSE, warning=FALSE}
#| label: tbl-all-trials
#| tbl-cap: "Summary of included RCTs"
#| tbl-cap-location: top

@@

 png("plot_means_multi_no_wl.png")
plot_means_no_wl
 dev.off()

@@

 png("plot_means_multi_cdrs.png")
 plot_means_cdrs
 dev.off()

@@

# subtitle_text_hamd <- "Metanalytic estimates of within-group changes: HAM-D studies only"
plot_means_multi_hamd <- plot_means_function(coef_and_se_means_hamd_study)
print(plot_means_multi_hamd)

 png("plot_means_multi_hamd.png")
 plot_means_multi_hamd
 dev.off()

@@

```{r, warning=FALSE, message = FALSE, echo=FALSE}
#| label: fig-plot-means-var2
#| fig-cap: "Meta-analytic estimates of within-group changes: studies with variance <0.02"
#| fig-cap-location: top
